Hydra battles in Coq

### 2020-08-31 latex/chapter-ks.tex
typo corrections in *.tex
new table of notations

\index{Functions!canonS canonS (helper for canon)}
\index{Functions!canon @ canon (canonical sequence)}
\begin{Coqsrc}
Fixpoint canonS alpha (i:nat) :=
  match alpha with
      zero => zero
    | ocons zero 0 zero => zero
    | ocons zero (S k) zero => FS k
    | ocons gamma 0 zero =>
      match pred gamma with
          Some gamma' => ocons gamma' i zero
        | None => ocons (canonS gamma i) 0 zero
      end
    | ocons gamma (S n) zero =>
       match pred gamma with
           Some gamma' => ocons gamma n (ocons gamma' i zero)
         | None => ocons gamma n (ocons (canonS gamma i) 0 zero)
       end
    | ocons alpha n beta => ocons alpha n (canonS beta i)
  end.
\end{Coqsrc}

@@

Let us consider a kind of accessibility problem inside $\epsilon_0$: given two ordinals $\alpha$ and $\beta$, where $\beta<\alpha<\epsilon_0$, find a \emph{path} consisting of a finite sequence $\gamma_0=\alpha,\dots,\gamma_l=\beta$,
where, for every $i<l$, $\gamma_i \not= 0$ and there exists some strictly positive integer $s_i$
such that $\gamma_{i+1}=\canonseq{\gamma}{s_i})$,

@@

\begin{exercise}[Continuation of exercise~\vref{exo:simply-typed-canonseq}]
Define a simply typed function for computing a path from $\alpha$ to $\beta$.
\end{exercise}

% ici
\noindent 
From the lemma \texttt{canonS\_LT}~\vref{lemma:canonS_LT}, we convert any path into an inequality on ordinals (by induction on paths).

>>>>

\index{Functions!canonS@canonS (helper for canon)}
\index{Functions!canon @canon (canonical sequence)}
\begin{Coqsrc}
Fixpoint canonS alpha (i:nat) :=
  match alpha with
      zero => zero
    | ocons zero 0 zero => zero
    | ocons zero (S k) zero => FS k
    | ocons gamma 0 zero =>
      match pred gamma with
          Some gamma' => ocons gamma' i zero
        | None => ocons (canonS gamma i) 0 zero
      end
    | ocons gamma (S n) zero =>
       match pred gamma with
           Some gamma' => ocons gamma n (ocons gamma' i zero)
         | None => ocons gamma n (ocons (canonS gamma i) 0 zero)
       end
    | ocons alpha n beta => ocons alpha n (canonS beta i)
  end.
\end{Coqsrc}

@@

Let us consider a kind of accessibility problem inside $\epsilon_0$: given two ordinals $\alpha$ and $\beta$, where $\beta<\alpha<\epsilon_0$, find a \emph{path} consisting of a finite sequence $\gamma_0=\alpha,\dots,\gamma_l=\beta$,
where, for every $i<l$, $\gamma_i \not= 0$ \footnote{This condition allows us to ignore paths which end by a lot of useless $0$.} and there exists some strictly positive integer $s_i$
such that $\gamma_{i+1}=\canonseq{\gamma}{s_i}$,

@@

\begin{exercise}[continuation of exercise~\vref{exo:simply-typed-canonseq}]
Define a simply typed function for computing a path from $\alpha$ to $\beta$.
\end{exercise}

% ici
\noindent 
From the lemma \texttt{canonS\_LT}~\vref{lemma:canonS_LT}, we can convert any path into an inequality on ordinals (by induction on paths).

### 2020-08-31 latex/chapter-ordinals.tex
corrections index
new table of notations

The following definitions allow us to make implicit several guessable arguments.

\begin{Coqsrc}
Definition on_t  {A:Type}{lt: relation A}
            {compare : A -> A -> comparison}
            {on : ON lt compare} := A.

@@

@@

It would be interesting to specify at the most generic level, what is a zero, a successor or a limit ordinal. Let $<$ be a strict order on a type $A$.

@@

Using these definitions , we can prove the following lemma:

@@

\index{Coq!Commands!Program Definition}

@@

Note also that attempting to compare an ordinal of type \texttt{t $n$}  with an ordinal of
type \texttt{t $p$}  leads to an error if $n$ and $p$ are not convertible.

>>>>

The following definitions allow us to make implicit several guessable arguments.
\label{sect:on-lt-notation}
\label{sect:on-le-notation}
\begin{Coqsrc}
Definition on_t  {A:Type}{lt: relation A}
            {compare : A -> A -> comparison}
            {on : ON lt compare} := A.

@@

\begin{remark}
The infix notations \texttt{o<} and \texttt{o<=} were defined in order to make apparent the distinction between the various notation scopes that may co-exist in a statement. In the informal and semi-formal statements, we still use $<$ and $\leq$.
\end{remark}

@@

\subsection{Definitions}
It would be interesting to specify at the most generic level, what is a zero, a successor or a limit ordinal. Let $<$ be a strict order on a type $A$.

@@

\index{Exercises}
\begin{exercise}
Prove, that, in any ordinal notation system, that if $\beta$ is a successor of $\alpha$,
then for any $\gamma$, $\gamma<\beta$ implies 
$\gamma\leq\alpha$.
\end{exercise}

\subsection{Limits and successors in \texorpdfstring{$\omega+\omega$}{omega+omega}}

Using thedefinitions above, we can prove the following lemma:

@@

@@

Note also that attempting to compare a term  of type (\texttt{t $n$}) with a term of
type (\texttt{t $p$})  leads to an error if $n$ and $p$ are not convertible.

### 2020-08-31 latex/chapter-termination.tex
minor corrections
corrections index
new table of notations

\begin{remark}
The proof of well-foundedness using \'Evelyne Contejean's work on recursive path ordering is available in the module \href{../src/html/hydras.Epsilon0.Epsilon0rpo.html}{Epsilon0rpo}.
 \end{remark}

@@

@@

\subsection{More theorems on Hessenberg's sum}

>>>>

\begin{remark}
The alternate proof of well-foundedness using \'Evelyne Contejean's work on recursive path ordering is available in the library \href{../src/html/hydras.Epsilon0.Epsilon0rpo.html}{Epsilon0rpo}.
 \end{remark}

@@

\label{sect:infix-oplus}

@@

\index{Projects}
\begin{project}
Compare \texttt{oplus}'s definition (with inner fixpoint) with other possibilities
(\texttt{coq-equations}, \texttt{Function}, etc.).
\end{project}
\subsection{More theorems on Hessenberg's sum}

### 2020-08-31 latex/hydras.tex
new chapter-skippable.tex
renaming zero_succ_limit-dec
simplification of class OrdinalNotation
corrections in the pdf documentation
correction on pdf file
proof simplification in L_alpha
added old Denumerable.v (to adapt)
new table of notations

\documentclass{book}
\usepackage{alltt}
\usepackage{makeidx}
\usepackage{hyperref}
\usepackage{url}
\usepackage{mdframed}
\usepackage{amsmath}
\usepackage{varioref}
\usepackage{synttree, proof, centernot}
\usepackage[firstpage]{draftwatermark}
\usepackage{verbatim}
\usepackage[note]{marginote}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathtools}
\usepackage{dsfont}
\setcounter{tocdepth}{1}
\input{coqmacros}

\author{Pierre Castéran\\ LaBRI, Univ. Bordeaux, CNRS UMR 5800, Bordeaux-INP}
\date{\today}
\title{The hydra and the rooster}
\makeindex
\begin{document}
\maketitle

@@

\include{introduction}
\include{chapter-hydras}
\include{chapter-termination}
\include{chapter-ks}
\include{chapter-large-sets}
\include{chapter-schutte}

@@

\bibliographystyle{alpha}
\bibliography{thebib, thebib2}

\section{Future work (projects)}
This document and the proof scripts are far from being complete.

@@

\section{Index}
\printindex

>>>>

\documentclass[a4paper]{book}
\usepackage{alltt}
\usepackage{makeidx}
\usepackage{hyperref}
\usepackage{url}
\usepackage{mdframed}
\usepackage{amsmath}
\usepackage{varioref}
\usepackage{synttree, proof, centernot}
\usepackage[firstpage]{draftwatermark}
\usepackage{verbatim}
\usepackage[note]{marginote}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathtools}
\usepackage{dsfont}
\setcounter{tocdepth}{2}
\input{coqmacros}

\author{Pierre Castéran\\ LaBRI, Univ. Bordeaux, CNRS UMR 5800, Bordeaux-INP\\with contributions by \'Evelyne Contejean and Florian Hatat}
\date{\today}
\title{Hydra battles in Coq}
\makeindex
\begin{document}
\maketitle

@@

\include{chapter-introduction}
\include{chapter-hydras}
\include{chapter-ordinals}
\include{chapter-termination}
\include{chapter-ks}
\include{chapter-large-sets}
\include{chapter-schutte}

@@

\bibliographystyle{alpha}
\bibliography{thebib}

\section{Future work (projects)}
\index{Projects}

This document and the proof scripts are far from being complete.

@@

\section{Index}
\subsection{Infix notations}

\begin{tabular}{|c|c|c|c|l|}
\hline
Name & Ascii&Math& Description& Where \\\hline
 \texttt{on\_lt} & \texttt{alpha o< beta}&$\alpha<\beta$& ordinal inequality & \vref{sect:on-lt-notation}\\
 \texttt{on\_le} & \texttt{alpha o<= beta}&$\alpha\leq\beta$& ordinal inequality & \vref{sect:on-lt-notation}\\
\texttt{oplus} & \texttt{alpha o+ beta} & $\alpha \oplus \beta$ & Hessenberg sum & \vref{sect:infix-oplus} \\
\texttt{round} & \texttt{h -1-> h'} & & round of a battle & \vref{sect:infix-round} \\
\hline
\end{tabular}

\printindex
